---
title: "Communication with DaVinci modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Communication with DaVinci modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Swimmer Plot module is capable of communicating with other DaVinci modules like the Patient Profile module from the [{dv.papo}](https://boehringer-ingelheim.github.io/dv.papo/) package. Communication in this sense means that the Swimmer Plot module sends out a subject ID which can be received and further processed by other modules. At the same moment, the tab of a DaVinci app switches to the receiver module.

The communication feature is optional and can be activated on the Swimmer Plot module side as follows. Note that there might be also activation needed on the counter module side (e.g., Patient Profile module side).

## Activate communication within Swimmer Plot module

As a default, the `receiver_id` parameter in the `mod_swimmerplot()` call of your module list definition is set to `NULL`, which means that the communication functionality is disabled. To enable switching to another DaVinci module, set `receiver_id` to the module ID of your counterpart module.

## Example

Example code of a module list definition to turn the communication feature on between a Swimmer Plot and a Patient Profile module:

```{r, eval=FALSE}
dm <- pharmaversesdtm::dm |>
  dplyr::select(USUBJID, AGE, SEX, RACE, ARM, RFSTDTC, RFENDTC) |> 
  dplyr::filter(ARM != "Screen Failure")

ex <- dplyr::left_join(x = pharmaversesdtm::ex, y = dm, by = "USUBJID") |>
  dplyr::mutate(
    ex_ongoing = is.na(EXENDY)
  )

# Example using oncology response data - the module works with any type of response or outcome data
rs <- dplyr::left_join(x = pharmaversesdtm::rs_onco, y = dm, by = "USUBJID") |>
  dplyr::filter(RSTEST == "Overall Response") |>
  dplyr::filter(RSEVAL == "INVESTIGATOR") |>
  dplyr::filter(!is.na(RSDY))

data_list <- list(dm = dm, ex = ex, rs = rs)

attributes(data_list$dm)$label <- "Demographics"
attributes(data_list$ex)$label <- "Exposure"
attributes(data_list$rs)$label <- "Response Assessment"

module_list <- list(
  "Swimmer Plot" = dv.swimmerplot::mod_swimmerplot(
    module_id = "swimmer1",
    subject_level_dataset_name = "dm",
    exposure_dataset_name = "ex",
    response_dataset_name = "rs",
    subjid_var = "USUBJID",
    group_by_vars = c("SEX", "RACE", "ARM"),
    sort_by_vars = c("AGE", "SEX", "RACE"),
    sort_direction = "asc",
    trt_start_day_var = "EXSTDY",
    trt_end_day_var = "EXENDY",
    trt_ongoing_var = "ex_ongoing",
    trt_tooltip_vars = c(
      "Subject ID: " = "USUBJID",
      "Exposure: " = "EXTRT",
      "Start Day: " = "EXSTDY",
      "End Day: " = "EXENDY"
    ),
    trt_group_var = "EXTRT",
    trt_legend_label = "Treatment",
    result_study_day_var = "RSDY",
    result_tooltip_vars = c(
      "Subject ID: " = "USUBJID",
      "Study Day: " = "RSDY",
      "Response: " = "RSORRES"
    ),
    result_cat_var = "RSORRES",
    result_legend_label = "Response",
    plot_title = "Subject Treatment and Response",
    plot_x_label = "Study Day",
    plot_y_label = "Subject ID",
    plot_width = 10,
    receiver_id = "papo1"
  ),
  "Patient Profile" = dv.papo::mod_patient_profile(
    module_id = "papo1",
    subject_level_dataset_name = "dm",
    subjid_var = "USUBJID",
    summary = list(
      vars = c("USUBJID", "AGE", "SEX", "RACE", "ARM"),
      column_count = 3L
    ),
    listings = list(
      "Exposure" = list(
        dataset = "ex",
        default_vars = c("EXSTDY", "EXENDY", "EXTRT", "EXDOSE", "EXDOSU")
      ),
      "Response" = list(
        dataset = "rs",
        default_vars = c("RSDY", "RSORRES", "RSTEST")
      )
    ),
    sender_ids = "swimmer1"
  )
)

dv.manager::run_app(
  data = list("Study Data" = data_list),
  module_list = module_list,
  filter_data = "dm",
  filter_key = "USUBJID"
)
```

For more information about module communication in DaVinci apps, see the [`dv.manager` documentation](https://boehringer-ingelheim.github.io/dv.manager/articles/passing_data.html).

For details on how to configure the Patient Profile module to receive communication, see the [`dv.papo` documentation on communication](https://boehringer-ingelheim.github.io/dv.papo/articles/a02-communication.html).
