---
title: "Module Customization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Module Customization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette covers the customization options available in `dv.swimmerplot`. Examples build on the data prepared in the [Swimmer Plot Module](dv-swimmerplot.html) vignette.

## Data Setup

```{r setup}
library(dv.swimmerplot)

dm <- pharmaversesdtm::dm |>
  dplyr::filter(ARM != "Screen Failure") |>
  dplyr::select(USUBJID, AGE, SEX, RACE, ARM, RFSTDTC, RFENDTC)

ex <- dplyr::left_join(pharmaversesdtm::ex, dm, by = "USUBJID") |>
  dplyr::mutate(
    ex_trt     = paste(EXTRT, EXDOSE, EXDOSU),
    ex_ongoing = is.na(EXENDY),
    ex_end     = ifelse(ex_ongoing, EXSTDY + 10, EXENDY)
  )

rs <- dplyr::left_join(pharmaversesdtm::rs_onco, dm, by = "USUBJID") |>
  dplyr::filter(RSTEST == "Overall Response", RSEVAL == "INVESTIGATOR", !is.na(RSDY))

sdtm_datasets <- list(dm = dm, ex = ex, rs = rs)
```

---

## Color Palette

Use `color_palette` to assign specific colors to exposure groups. Supply a **named character vector** where names match values of `trt_group_var`.

```{r color-palette}
mod_swimmerplot(
  module_id                = "swimmer_colors",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  response_dataset_name    = "rs",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  result_study_day_var     = "RSDY",
  result_cat_var           = "RSORRES",
  color_palette = c(
    "PLACEBO 0 mg"       = "#AAAAAA",   # grey for placebo
    "XANOMELINE 54 mg"   = "#4E79A7",   # blue
    "XANOMELINE 81 mg"   = "#F28E2B"    # orange
  )
)
```

When `color_palette` is `NULL` (the default), ggplot2 automatically assigns colors.

---

## Shape Mapping for Response Points

Use `shape_mapping` to control the point shape for each response category. Supply a **named numeric vector** where names match values of `result_cat_var` and values are R `pch` codes.

```{r shape-mapping}
mod_swimmerplot(
  module_id                = "swimmer_shapes",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  response_dataset_name    = "rs",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  result_study_day_var     = "RSDY",
  result_cat_var           = "RSORRES",
  shape_mapping = c(
    "CR" = 16,   # filled circle  — Complete Response
    "PR" = 17,   # filled triangle — Partial Response
    "SD" = 15,   # filled square  — Stable Disease
    "PD" = 18    # filled diamond — Progressive Disease
  )
)
```

---

## Tooltips

Both exposure bars and response points support interactive tooltips. Pass a **named character vector** where names are the label text and values are column names. Labels appear before the value in the tooltip (e.g., `"Dose: 54"`).

```{r tooltips}
mod_swimmerplot(
  module_id                = "swimmer_tooltips",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  response_dataset_name    = "rs",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  trt_tooltip_vars = c(
    "Subject: "   = "USUBJID",
    "Treatment: " = "ex_trt",
    "Start Day: " = "EXSTDY",
    "End Day: "   = "EXENDY"
  ),
  result_study_day_var = "RSDY",
  result_cat_var       = "RSORRES",
  result_tooltip_vars = c(
    "Subject: "   = "USUBJID",
    "Study Day: " = "RSDY",
    "Response: "  = "RSORRES"
  )
)
```

Omit `trt_tooltip_vars` or `result_tooltip_vars` (or set them to `NULL`) to disable tooltips for that layer.

---

## Bar Annotations

`trt_annotation_vars` adds text labels directly on the plot, next to the end of each subject's last exposure bar. This is useful for annotating treatment summaries or end-of-study status.

```{r annotations-trailing}
# Annotation placed at the end of each subject's last exposure bar (default)
mod_swimmerplot(
  module_id                = "swimmer_annot_trail",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  trt_annotation_vars      = c("ARM"),   # column(s) to concatenate as annotation text
  trt_annotation_x         = NULL        # NULL = place annotation after bar end
)
```

Use `trt_annotation_x` to pin all annotations to a fixed x position, creating an aligned column:

```{r annotations-fixed}
mod_swimmerplot(
  module_id                = "swimmer_annot_fixed",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  trt_annotation_vars      = c("ARM"),
  trt_annotation_x         = 200  # fixed x position for all annotation labels
)
```

---

## Sorting

Control subject ordering with `sort_by_vars` and `sort_direction`. Multiple variables are supported and applied in order.

```{r sorting}
mod_swimmerplot(
  module_id                = "swimmer_sort",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  sort_by_vars             = c("ARM", "AGE"),  # primary sort: ARM, secondary: AGE
  sort_direction           = "desc"            # "asc" or "desc"
)
```

Users can also change sorting interactively at runtime via the **Plot Options** dropdown. The `sort_by_vars` and `sort_direction` arguments set the initial defaults.

---

## Grouping (Faceting)

`group_by_vars` splits the plot into facets by one or more categorical variables. Each combination gets its own panel with a free y-axis scale.

```{r grouping}
mod_swimmerplot(
  module_id                = "swimmer_group",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  group_by_vars            = c("SEX", "RACE")  # facet by sex and race
)
```

Set `group_by_vars = NULL` to display all subjects in a single unfaceted panel.

---

## Filter Control

The module includes a built-in filter dropdown. Use `filter_var` to choose which variable drives the filter and `filter_values` to restrict the selectable choices.

```{r filter-basic}
mod_swimmerplot(
  module_id                = "swimmer_filter",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  response_dataset_name    = "rs",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  result_study_day_var     = "RSDY",
  result_cat_var           = "RSORRES",
  filter_var               = "ARM",       # variable to filter on
  filter_values            = c("Xanomeline High Dose", "Xanomeline Low Dose")
)
```

### Filtering from a Different Dataset

By default the filter is drawn from the subject-level dataset. Set `filter_data` to the name of any dataset in your app (e.g. the response dataset) to populate filter choices from there instead:

```{r filter-source}
mod_swimmerplot(
  module_id                = "swimmer_filter_rs",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  response_dataset_name    = "rs",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  result_study_day_var     = "RSDY",
  result_cat_var           = "RSORRES",
  filter_data              = "rs",           # draw filter choices from rs
  filter_var               = "RSORRES",      # filter variable in rs
  filter_values            = c("CR", "SD"),  # restrict available choices
  filter_default_vals      = c("CR", "SD")   # pre-select these on startup
)
```

---

## Axis Labels and Plot Titles

All text labels on the plot are configurable:

```{r labels}
mod_swimmerplot(
  module_id                = "swimmer_labels",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  trt_legend_label         = "Treatment Arm",
  plot_title               = "Subject-Level Exposure Timeline",
  plot_subtitle            = "Arrows indicate ongoing treatment",
  plot_x_label             = "Days from First Dose",
  plot_y_label             = "Patient ID"
)
```

---

## Plot Dimensions

Control the initial plot size (in inches). `plot_height = NULL` auto-scales height based on the number of subjects (approximately 0.3 inches per subject, minimum 6 inches).

```{r dimensions}
mod_swimmerplot(
  module_id                = "swimmer_size",
  subject_level_dataset_name = "dm",
  exposure_dataset_name    = "ex",
  subjid_var               = "USUBJID",
  trt_start_day_var        = "EXSTDY",
  trt_end_day_var          = "ex_end",
  trt_group_var            = "ex_trt",
  trt_ongoing_var          = "ex_ongoing",
  plot_width               = 14,   # width in inches
  plot_height              = NULL  # auto-calculated from subject count
)
```

---
